---
title: "HW11 Data Visualization II"
author: "Andrea Winingar"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

Load packages.

```{r}
library(tidyverse)
library(here)
library(GGally)
```

### Data exploration: Crabs

Import data, then remove some columns, remove NA, create row_order dummy variable

```{r}
import_crabs <- read_csv(here('data', 'aegla_crabs.csv'))
crabs_data <- select(import_crabs, Width, Depth, Flow, pH, TDS, Cond, N, Ni, Nt, Phos, Mg, Ca)
crabs_no_na <- drop_na(crabs_data, N, Ni, Nt, Phos, Mg, Ca)
crabs <-  crabs_no_na %>% 
  mutate(box_group = 'Crabs',
         row_order = 1:nrow(.))
```

#### Identifying outlier

Boxplots to identfy outliers.

```{r}
crabs %>% 
  ggplot(aes(x = box_group, 
             y = Width)) +
  geom_boxplot(width = 0.25)
crabs %>% 
  ggplot(aes(x = box_group, 
             y = Depth)) +
  geom_boxplot(width = 0.25)
crabs %>% 
  ggplot(aes(x = box_group, 
             y = Flow)) +
  geom_boxplot(width = 0.25)
crabs %>% 
  ggplot(aes(x = box_group, 
             y = pH)) +
  geom_boxplot(width = 0.25)
crabs %>% 
  ggplot(aes(x = box_group,
             y = TDS)) +
  geom_boxplot(width = 0.25)
crabs %>% 
  ggplot(aes(x = box_group, 
             y = Cond)) +
  geom_boxplot(width = 0.25)
crabs %>% 
  ggplot(aes(x = box_group, 
             y = N)) +
  geom_boxplot(width = 0.25)
crabs %>% 
  ggplot(aes(x = box_group, 
             y = Ni)) +
  geom_boxplot(width = 0.25)
crabs %>% 
  ggplot(aes(x = box_group, 
             y = Nt)) +
  geom_boxplot(width = 0.25)
crabs %>% 
  ggplot(aes(x = box_group, 
             y = Phos)) +
  geom_boxplot(width = 0.25)
crabs %>% 
  ggplot(aes(x = box_group, 
             y = Mg)) +
  geom_boxplot(width = 0.25)
crabs %>% 
  ggplot(aes(x = box_group, 
             y = Ca)) +
  geom_boxplot(width = 0.25)
```

Phos, Depth and Flow all have outliers, outside the interquartile range by 1.5*. I believe Phosphate has the most outlieing outlier. 


#### Cleveland Plots

Make four different cleveland plots:

```{r}
crabs %>% 
  ggplot(aes(x = Width,
             y = row_order)) +
  geom_point() +
  labs(y = "Order of the data")
crabs %>% 
  ggplot(aes(x = Depth,
             y = row_order)) +
  geom_point() +
  labs(y = "Order of the data")
crabs %>% 
  ggplot(aes(x = Flow,
             y = row_order)) +
  geom_point() +
  labs(y = "Order of the data")
crabs %>% 
  ggplot(aes(x = pH,
             y = row_order)) +
  geom_point() +
  labs(y = "Order of the data")
```

Width- This graph shows the width being in a large range that agrees with the boxplot previously made of the interquartile range. Depth- Typically shows a range between 0.2 and 2.5 with the outlier. Flow- has a range of 0.2 to 0.55 with a few variables including the outlier outside of the main range. pH- has a rather neutral pH overall, a few that could be more on  basic/acidic side.

#### Histograms

Make three histograms and save to a variable. adjust bins argument

```{r}
h1 <- crabs %>% 
  ggplot() +
  geom_histogram(aes(x = N),
                 color = 'black',
                 bins = 15,
                 boundary = 0,
                 fill = 'blue')
h2 <- crabs %>% 
  ggplot() +
  geom_histogram(aes(x = Ni),
                 color = 'black',
                 bins = 25,
                 boundary = 0,
                 fill = 'green')
h3 <- crabs %>% 
  ggplot() +
  geom_histogram(aes(x = Nt),
                 color = 'black',
                 bins = 20,
                 boundary = 0,
                 fill = 'purple')
h1
```

I probably had way too much fun messing around and adjusting these histograms. The bins() changed the width of the individual columns or bars in the histograph. The N and Nt concentrations are on a much smaller scale than the Ni concentrations, meaning that there is more Ni overal than N or Nt.






