---
title: "HW05: Data Visulaization"
author: "Andrea Winingar"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# First Things firt, load the Tidyverse.

```{r}
library('tidyverse')
```

#### Plot 1

Load the dataset.

```{r}
data(trees)
?trees
```


Now to create the scatterplot.

```{r}
ggplot(data = trees) +
  geom_point(mapping = aes(x = Height, y = Girth))
```

The larger the Height of the tree, the larger the girth of the tree is. 

#### Plot 2

Make a vector.

```{r}
year <- c(1821:1934)
```

Look at class() of 'lynx' dataset.

```{r}
class(lynx)
```

Change lynx from 'ts' to a vector.

```{r}
pelts <- as.vector(lynx)
```

Make a dataframe with the vectors.

```{r}
lynx_pelts <- data.frame('year' = 1821:1934, 'pelts' = as.vector(lynx)) 
```

Make the line graph plot.

```{r}
ggplot(data = lynx_pelts) +
  geom_line(mapping = aes(x = year, y = pelts), color = "maroon")
```

The graph is kind of wavey. It goes up and down a lot.

#### Plot 3

load the dataset irirs.

```{r}
data("iris")
?iris
```

Make the scatterplot

```{r}
ggplot(data = iris) +
  geom_point(mapping = aes(x = Petal.Length, y = Petal.Width, color = Species, shape = Species), size = 2) + 
  labs(x = 'Petal Length', y = 'Petal Width')

```

Each species is seperate from the others, they are bunched together. If the petal is longer, it is also wider.

#### Plots 4 and 5

Load the dataset

```{r}
data("msleep")
```

This code chunk is Plot 4.

```{r}
ggplot(data = msleep) +
  geom_violin(mapping = aes(x = vore, y = sleep_rem), fill = "gray30")
```

This plot varies, some of the vore's have lower ranges of sleep_rem and the others have their range in higer sleep_rem.

This code chunk is plot 5.

```{r}
ggplot(data = msleep) +
  geom_violin(mapping = aes(x = vore, y = sleep_rem), fill = "gray70")
```

This is the same as the other violin plot but a different gray. The violin plot shows the same as box plot does with median and interquartile ranges, but violin plots also shows the full distribution of data.