---
title: "HW 08: More Data Wrangling"
author: "Andrea WIningar"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- Change "your name" in the YAML header above to your name.

- As usual, enter the examples in code chunks and run them, unless told otherwise.

## Chapter 5: Data transformation

Read [R4ds Chapter 5: Data transformation](https://r4ds.had.co.nz/transform.html), sections 5-7.

**You do not answer any of the questions this week.** Just read and run all of the examples as noted below. Pay attention to the code you enter for the examples. You will be applying the information in future assignments.

### Section 5.5: Add new variables with `mutate()`

Type and run all of the examples.

Pay careful attention to the useful creation functions in section 5.5.1. 

This code chunk loads flights data, and tidyverse.

```{r}
library(nycflights13)
library(tidyverse)
```


This code chunk creates a narower dataset to use in the next code chunk.

```{r}
flights_sml <- select(flights, 
  year:day, 
  ends_with("delay"), 
  distance, 
  air_time
)
```

This code chunk uses mutate() to add new columns at the end of the dataset.

```{r}
mutate(flights_sml,
  gain = dep_delay - arr_delay,
  speed = distance / air_time * 60
)
```

This code chunk shows you can refer to the columns just created in chunk before.

```{r}
mutate(flights_sml,
  gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours
)
```

This code chunk uses transmute() to keep only the new variables.

```{r}
transmute(flights,
  gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours
)
```

This code chunk shows how to compute hour and minute from dep_time.

```{r}
transmute(flights,
  dep_time,
  hour = dep_time %/% 100,
  minute = dep_time %% 100
)
```

This code chunk shows examples of lead() and lag()

```{r}
(x <- 1:10)
lag(x)
lead(x)
```

This code chunk shows how to use cumsum() and cummean()

```{r}
x
cumsum(x)
cummean(x)
```

This code chunk shows how to use min_rank().

```{r}
y <- c(1, 2, 2, NA, 3, 4)
min_rank(y)
min_rank(desc(y))
```

This code chunk shows how to use row_number(), dense_rank(), percent_rank(), cume_dist().

```{r}
row_number(y)
dense_rank(y)
percent_rank(y)
cume_dist(y)
```

### Section 5.6: Grouped summaries with `summarise()`

Read and run all of the examples. Pay careful attention to the useful summary functions. Unless you really like baseball, skip the baseball part of 5.6.3.

This code chunk shows how summarise() collapses a data frame to a single row.

```{r}
summarise(flights, delay = mean(dep_delay, na.rm = TRUE))
```

This code chunk uses summarise() paired with group_by().

```{r}
by_day <- group_by(flights, year, month, day)
summarise(by_day, delay = mean(dep_delay, na.rm = TRUE))
```

This code chunk shows the relationship between distance and average delay.

```{r}
by_dest <- group_by(flights, dest)
delay <- summarise(by_dest,
  count = n(),
  dist = mean(distance, na.rm = TRUE),
  delay = mean(arr_delay, na.rm = TRUE)
)
delay <- filter(delay, count > 20, dest != "HNL")
ggplot(data = delay, mapping = aes(x = dist, y = delay)) +
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)
```

This code chunk tackles the same as the on above but with use of the pipe.

```{r}
delays <- flights %>% 
  group_by(dest) %>% 
  summarise(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>% 
  filter(count > 20, dest != "HNL")
```

This code chunk shows what happens without na.rm

```{r}
flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))
```

This code chunk includes na.rm to remove na

```{r}
flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay, na.rm = TRUE))
```

This code chunk removes canceled flights.

```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))
```

This code chunk looks at the planes with the highest average delays.

```{r}
delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay)
  )
ggplot(data = delays, mapping = aes(x = delay)) + 
  geom_freqpoly(binwidth = 10)
```

This code chunk gives more insight with a scatterplot of number of lights against average delay.

```{r}
delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay, na.rm = TRUE),
    n = n()
  )
ggplot(data = delays, mapping = aes(x = n, y = delay)) + 
  geom_point(alpha = 1/10)
```

This code chunk filters out the groups with the smallest numbers of observations.

```{r}
delays %>% 
  filter(n > 25) %>% 
  ggplot(mapping = aes(x = n, y = delay)) + 
    geom_point(alpha = 1/10)
```

This code chunk combines aggregation with logical subsetting.

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    avg_delay1 = mean(arr_delay),
    avg_delay2 = mean(arr_delay[arr_delay > 0]) # the average positive delay
  )
```

This code chunk shows why the distance to some destinations more varaible than to others with sd()

```{r}
not_cancelled %>% 
  group_by(dest) %>% 
  summarise(distance_sd = sd(distance)) %>% 
  arrange(desc(distance_sd))
```

This code chunk shows when the first and last flights leave each day with min() and max()

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    first = min(dep_time),
    last = max(dep_time)
  )
```

This code chunk finds the first and last departure for each day with first() and last()

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    first_dep = first(dep_time), 
    last_dep = last(dep_time)
  )
```

This code chunk shows the functions that are complimentary to filtering on ranks, gives you all variables, with each observation in sepearte row.

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  mutate(r = min_rank(desc(dep_time))) %>% 
  filter(r %in% range(r))
```

This code chunk shows what destinations have the most carriers, while using n_dixtinct() to count the number of unique variables.

```{r}
not_cancelled %>% 
  group_by(dest) %>% 
  summarise(carriers = n_distinct(carrier)) %>% 
  arrange(desc(carriers))
```

This code chunk shows the helper dplyr provides if all you want is a count.

```{r}
not_cancelled %>% 
  count(dest)
```

This code chunk shows how to optionally apply a weight variable.

```{r}
not_cancelled %>% 
  count(tailnum, wt = distance)
```

This code chun shows how many flights left before 5 am. Then what proportion of flights are delayedby more than an hour.

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(n_early = sum(dep_time < 500))
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(hour_prop = mean(arr_delay > 60))
```

This code chunk shows how to group  by multiple varaibles, making it easier to roll up a dataset.

```{r}
daily <- group_by(flights, year, month, day)
(per_day   <- summarise(daily, flights = n()))
(per_month <- summarise(per_day, flights = sum(flights)))
(per_year  <- summarise(per_month, flights = sum(flights)))
```

This code chunk uses ungroup() to remove grouping.

```{r}
daily %>% 
  ungroup() %>% summarise(flights = n())
```

### Section 5.7: Grouped mutates (and filters)

Read and run all of the examples.