---
title: "HW 07: Data Importing and Wrangling I"
author: "Andrea Winingar"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

Load the packages and make link to data.

```{r}
library(tidyverse)
library(here)
```

### Limpets

```{r}
limpets_data <- read_csv(here('data','limpets.csv'))
```

The data is tidy.

using col_types argument.

```{r}
limpets_data <- read_csv("limpets.csv", col_types = "ncc")
```

Factor Site column.

```{r}
factor(limpets_data$Site, levels = c('PBL', 'KNRM', 'VBG', 'WP', 'PF', 'DP', 'SIO', 'CTZ1', 'CTZ2', 'CNM'))
```

Boxplot of limpets_data

```{r}
ggplot(data = limpets_data) +
  geom_boxplot(mapping = aes(x = Site, y = Length, color = Protected)) +
  labs(x = "Site", y = "Length (mm)")
```

DP and WP have outliers. The protected sites have larger limpets than nonprotected sites.The protected sites also have a larger box range. 

### Roseate Terns

Load the data.

```{r}
roseate_data <- read_tsv(here("data", "roseate_terns.txt"))
```

The data is not tidy. This next code chunk will tidy it.

```{r}
roseate_tidy <- roseate_data %>% 
  pivot_longer(`AR`:`JC`,
               names_to = "Site",
               values_to = "Population")
```

Use filter to remove sites with missing counts.

```{r}
filter(drop_na(roseate_tidy))
```

Make Lineplot of pop size over time.    

```{r}
roseate_tidy %>% 
  ggplot() +
  geom_line(mapping = aes(x = year, y = Population, color = Site)) +
  labs(x = 'Year', y = 'Population')
```

Sites BC, JC, AR all increased in population, AR increased the most dramatically. JR, BR both decreased drastically and all others did not have any significant changes. The non continuos ones could be caused by the population being NA and was removed.

### Blacklip Abalone

Chunk 1:  Import, remove the first column, then make a boxplot of height differences among the three types.

```{r}
abalone_data <- read_csv(here('data', 'abalone.csv'))
select(abalone_data, -('X1'))
abalone_data %>% 
  ggplot() +
  geom_boxplot(mapping = aes(x = Type, y = Height))
```

There are two outliers, one in the F and one in M.

Chunk 2: scatterplot to see if height appears to correlate with rings. 

```{r}
abalone_data %>% 
  ggplot() +
  geom_point(mapping = aes(x = Rings, y = Height))
```

Height correlates to Rings to an extent, but no enough to explain or cause the outliers.

Chunk 3:remove outlier by filtering. Make Type an ordered factor. Redo scatterplot.

```{r}
abalone_filtered <- filter(abalone_data, Height <= 0.4)
factor(abalone_filtered$Type, level = 'I', 'F', 'M')
abalone_filtered %>% 
  ggplot() +
  geom_point(mapping = aes(x = Rings, y = Height, color = Type))
```

Immatures are generally smaller, males and females are mixed together without either being distinguishably smaller or larger. There are some male and females that are smaller than the immatures.

Chunk 4: print records of those with zero height and change to NA. then replot.

```{r}
abalone_filtered[abalone_filtered$Height == 0, 'Height'] <- NA
abalone_filtered %>% 
  ggplot() +
  geom_point(mapping = aes(x = Rings, y = Height, color = Type))
```

Chunk 5: Make two scatterplots of choice.

```{r}
abalone_filtered %>% 
  ggplot() +
  geom_point(mapping = aes(x = Diameter, y = Height, color = Type))
abalone_filtered %>% 
  ggplot() +
  geom_point(mapping = aes(x = Rings, y = Diameter, color = Type))
```

The first scatterplot shows a relation of Immatures typically having the smallest diameter, then males and females are mixed together. Also, the larger the diameter the taller. The second scatterplot shows immature having the smallest diameter and fewest rings with a few males mixed in with them. The more rings, the larger the diameter.

### Darters

Load the data.

```{r}
darters_data <- read_fwf(here('data', 'darter.txt'), fwf_widths(c(3, 11, 1, 1, 1, 3, 4, 4, 4, 3, 3, 2, 2)), comment = '#')
```

Chunk 1: Make column titles, make riffle an unordered factor, and make major_type an ordered factor.

```{r}
darter_renamed <- darters_data %>% 
  rename(
    ID = X1,
    Species = X2,
    Riffle = X3,
    Age = X4,
    Sex = X5,
    Length = X6,
    Depth = X7,
    Velocity = X8,
    Temperature = X9,
    Maj_Sub = X10,
    Min_Sub = X11,
    Maj_Type = X12,
    Min_Type = X13
  )
darter_renamed$Riffle<- factor(darter_renamed$Riffle, levels = c('1', '2'))
darter_renamed$Maj_Type <- factor(darter_renamed$Maj_Type, levels = c('s', 'fg', 'sg', 'lg', 'c'), ordered = TRUE)
```

Chunk 2: Use filter to remove zonale and tetrazonum. Remove columns min_sub and min_type. Rename maj_sub and maj_type. arrange by id.

```{r}
darter_renamed %>% 
  filter(Species == "blennioides" | Species == "caeruleum" | Species == "spectabile") %>% 
  select(ID:Maj_Sub, Maj_Type) %>% 
  rename(
    major_substrate = Maj_Sub,
    major_type = Maj_Type) %>% 
  arrange(ID)
```

Plot 1: length as a function of depth, species to color and shape, facet_wrap() on riffles.

```{r}
darter_renamed %>% 
  ggplot() +
  geom_point(mapping = aes(x = Depth, y = Length, color = Species, Shape = Species)) +
  facet_wrap(~ Riffle, nrow = 1)
```

Riffle 2 has longer variables and riffle one has higher depths.

Plot 2: boxplot

```{r}
darter_renamed %>% 
  ggplot() +
  geom_boxplot(mapping = aes(x = Species, y = Length)) +
  facet_wrap(~ Riffle, nrow = 1)
```

Riffle two has several outliers, 