---
title: "HW 07: Data Importing and Wrangling I"
author: "Andrea Winingar"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

Load the packages and make link to data.

```{r}
library(tidyverse)
file_path <- "../data"
```

### Limpets

```{r}
limpets_data <- read_csv("limpets.csv")
```

The data is tidy.

using col_types argument.

```{r}
limpets_data <- read_csv("limpets.csv", col_types = "ncc")
```

Factor Site column.

```{r}
factor(limpets_data$Site, levels = c('PBL', 'KNRM', 'VBG', 'WP', 'PF', 'DP', 'SIO', 'CTZ1', 'CTZ2', 'CNM'))
```

Boxplot of limpets_data

```{r}
ggplot(data = limpets_data) +
  geom_boxplot(mapping = aes(x = Site, y = Length, color = Protected)) +
  labs(x = "Site", y = "Length (mm)")
```

DP and WP have outliers. The protected sites have larger limpets than nonprotected sites.The protected sites also have a larger box range. 

### Roseate Terns

Load the data.

```{r}
roseate_data <- read.csv("roseate_terns.txt")
```

The data is not tidy.

Use filter to remove sites with missing counts.

```{r}
filter(roseate_data, )
```



### Blacklip Abalone

Chunk 1:  Import, remove the first column, then make a boxplot of height differences among the three types.

```{r}
abalone_data <- read_csv('abalone.csv')
select(abalone_data, -('X1'))
abalone_data %>% 
  ggplot() +
  geom_boxplot(mapping = aes(x = Type, y = Height))
```
Chunk 2: scatterplot to see if height appears to correlate with rings. 

```{r}
abalone_data %>% 
  ggplot() +
  geom_point(mapping = aes(x = Rings, y = Height))
```

Height correlates to Rings to an extent, but no enough to explain or cause the outliers.

Chunk 3:

```{r}

```



### Darters


