---
title: "HW 07: Data Importing and Wrangling I"
author: "Andrea Winingar"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

Load the packages and make link to data.

```{r}
library(tidyverse)
library(here)
```

### Limpets

```{r}
limpets_data <- read_csv(here('data','limpets.csv'))
```

The data is tidy.

using col_types argument.

```{r}
limpets_data <- read_csv("limpets.csv", col_types = "ncc")
```

Factor Site column.

```{r}
factor(limpets_data$Site, levels = c('PBL', 'KNRM', 'VBG', 'WP', 'PF', 'DP', 'SIO', 'CTZ1', 'CTZ2', 'CNM'))
```

Boxplot of limpets_data

```{r}
ggplot(data = limpets_data) +
  geom_boxplot(mapping = aes(x = Site, y = Length, color = Protected)) +
  labs(x = "Site", y = "Length (mm)")
```

DP and WP have outliers. The protected sites have larger limpets than nonprotected sites.The protected sites also have a larger box range. 

### Roseate Terns

Load the data.

```{r}
roseate_data <- read_tsv(here("data", "roseate_terns.txt"))
```

The data is not tidy. This next code chunk will tidy it.

```{r}
roseate_tidy <- roseate_data %>% 
  pivot_longer(`AR`:`JC`,
               names_to = "Site",
               values_to = "Population")
```

Use filter to remove sites with missing counts.

```{r}
filter(drop_na(roseate_tidy))
```

Make Lineplot of pop size over time.    

```{r}
roseate_tidy %>% 
  ggplot() +
  geom_line(mapping = aes(x = year, y = Population, color = Site)) +
  labs(x = 'Year', y = 'Population')
```

Sites BC, JC, AR all increased in population, AR increased the most dramatically. JR, BR both decreased drastically and all others did not have any significant changes. The non continuos ones could be caused by the population being NA and was removed.

### Blacklip Abalone

Chunk 1:  Import, remove the first column, then make a boxplot of height differences among the three types.

```{r}
abalone_data <- read_csv(here('data', 'abalone.csv'))
select(abalone_data, -('X1'))
abalone_data %>% 
  ggplot() +
  geom_boxplot(mapping = aes(x = Type, y = Height))
```

There are two outliers, one in the F and one in M.

Chunk 2: scatterplot to see if height appears to correlate with rings. 

```{r}
abalone_data %>% 
  ggplot() +
  geom_point(mapping = aes(x = Rings, y = Height))
```

Height correlates to Rings to an extent, but no enough to explain or cause the outliers.

Chunk 3:remove outlier by filtering. Make Type an ordered factor. Redo scatterplot.

```{r}
filter(abalone_data, Height <= 0.4)
factor(abalone_data$Type, level = 'I', 'F', 'M')
abalone_data %>% 
  ggplot() +
  geom_point(mapping = aes(x = Rings, y = Height))
```



### Darters


