---
title: "HW09: Wrangling II"
author: "Andrea Winingar"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

Load packages.

```{r}
library(tidyverse)
library(here)
library(knitr)
library(smatr)
```

source functions.

```{r}
my_functions <- source(here('Scripts', 'my_functions.R'))
```

### Darters in Riffles

#### Import Data

This code chunk imports the darter data. Then use filter() to remove tetrazonum and zonale. 

```{r}
import_darter <- read_csv(here('data', 'darters.csv')) %>% 
  filter(species == "blennioides" | species == "caeruleum" | species == "spectabile")
```

#### Wrangle Data

Use mutate() to sum majsub and minsub in new column called total_substrate, change riffle value 1 and 2 to riffle 1 and riffle 2. change length from cm to mm. change sex values from f and m to female and male. 

```{r}
darter_data <- import_darter %>% 
  mutate(sex = ifelse(sex == "f", 
                             "female", 
                             "male"),
         total_substrate = majsub + minsub,
         length = length * 10,
         riffle = ifelse(riffle == "1",
                         "riffle 1",
                         "riffle 2"))
darter_data
```

#### Summarize Data

Summarize data with group_by and summarize. summarize mean length, depth, and velocity for each species for each riffle. 

```{r}
summarized_darter <- group_by(darter_data, riffle, species)
summarized_darter <- summarise(summarized_darter, 
          length = mean(length, na.rm = TRUE), 
          depth = mean(depth, na.rm = TRUE),
          velocity = mean(velocity, na.rm = TRUE))
summarized_darter
```

#### Table results

Use kable() to make table with summary means.

```{r}
summarized_darter %>% 
  kable()
```

#### Graph Data

Plot 1: uses stat_summary() technique x = species, y = length. add geom_point layer to add mean length of each species to the plot with filled square shape.

```{r}
ggplot(data = darter_data) +
  stat_summary(
    mapping = aes(x = species, y = length),
    fun = median,
    fun.min = min,
    fun.max = max
  ) +
  coord_flip() +
  facet_wrap(~ riffle) +
  labs( x = NULL, y = 'Length (mm)') +
  geom_point(data = summarized_darter, mapping = aes(x = species, y = length), shape = 15)
```

Plot 2: New plot with new geom_point layer of the length variable for each species. Change point color to gray60. add stat_summary() layer with arguments fun.y = median, geom = 'point', and size = 2.2. Add another stat_summary() change y function to mean, size to 3, and add color = 'maroon'. facet, flip and label. 

```{r}
darter_data %>% 
  ggplot() +
  geom_point(mapping = aes(x = species, y = length), color = 'gray60') +
  stat_summary(
    mapping = aes(x = species, y = length),
    fun.y = median,
    geom = 'point',
    size = 2.2
  ) +
  stat_summary(
    mapping = aes(x = species, y = length),
    fun.y = mean,
    size = 3,
    color = "maroon"
  ) +
  facet_wrap(~ riffle) +
  coord_flip() +
  labs(x = NULL, y = 'Length (mm)')
```


### Stress and Corticosterones in birds

#### Import data

This code chunk imports data and makes necessary changes. 

```{r}
import_feathers <- read_csv(here('data', 'feathers.csv'), locale = locale(decimal_mark = ','))
import_feathers
renamed_feathers <- rename(import_feathers,
                           capture_date = 'Capture Date',
                           day = 'Date',
                           site = 'Site',
                           urban_score = 'Urba Score',
                           cortF = 'CortF ng/mg',
                           cortB = 'CortF ng/mm',
                           sex = 'Sex',
                           smi = 'SMI',
                           tarsus = 'Tarsus',
                           mass = 'Mass',
                           basal_court = 'Bas_Cort',
                           stress_court = 'Stress_Cort'
)
```

#### Wrangle data

Calculate mean tarsus and save to mean_tarsus, use mutate() to log-transform(log()) the body mass and the tarsus length save in log-mass and log_tarsus. 

```{r}
mean_tarsus <- mean(renamed_feathers$tarsus, na.rm = TRUE)
log_mass <- mutate(renamed_feathers, log(mass))
log_tarsus <- mutate(renamed_feathers, log(tarsus))
mutated_feathers <- mutate(renamed_feathers,
       new_smi = scaled_mass(mass = renamed_feathers$mass,
                             tarsus = renamed_feathers$tarsus,
                             slope = 1.8))
```

#### Summarize Data

group by site and then use summarize() to calculate mean and standard error of the mean for basal_court and stress_court for each site. 

```{r}
grouped_feathers <- group_by(mutated_feathers, site)
summarized_feathers <- summarise(grouped_feathers,
                                 mean_basal = mean(basal_court, na.rm = TRUE),
                                 mean_stress = mean(stress_court, na.rm = TRUE),
                                 std_err_basal = std_err(basal_court, na.rm = TRUE),
                                 std_err_stress = std_err(stress_court, na.rm = TRUE))
summarized_feathers
```

#### Graph Data

Plot 1: make boxplot of new_smi for each site

```{r}
mutated_feathers %>% 
  ggplot() +
  geom_boxplot(mapping = aes(y = new_smi, x = site))
```

Each site has the box range after 23 smi, but there are no significant differences.

Plot 2: Make  plot of basal_court for each site.

```{r}
summarized_feathers %>% 
  ggplot() +
  geom_point((mapping = aes(y = std_err_basal, x = site))) +
  geom_errorbar(mapping = aes(ymin = std_err_basal - 1, ymax = std_err_basal + 1, x = site))
```





