---
title: "HW13: Strings and Things"
author: "Andrea Winingar"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

Load packages.

```{r}
library(tidyverse)
library(here)
library(ggthemes)
library(gghighlight)
```

### Part 1: DNA Strings

#### Import the Data

```{r}
beta_import <- read_lines(here('data', 'beta_gliadin.txt'), skip = 1)
print(beta_import)
```

#### From many Strings to One

Replace all whitespaces and digits with nothing. 

```{r}
beta_replaced <- beta_import %>% 
  str_replace_all("\\s", "") %>% 
  str_replace_all("\\d", "") %>% 
  str_c(collapse = "") %>% 
  str_to_upper()
beta_replaced
str_length(beta_replaced)
```

#### Sequence analysis: find the promoter region

Find how many possible start and stop codons.

```{r}
str_count(beta_replaced, "[^ACTG]")
str_count(beta_replaced, 'ATG')
str_count(beta_replaced, 'TGA|TAA')
```

Create caat_box and tata_box, count how many times they occur

```{r}
caat_box <- "CCAAT"
tata_box <- "CTATAA"
str_count(beta_replaced, caat_box)
str_count(beta_replaced, tata_box)
```

Create caat_seq anda tata_seq then count them.

```{r}
caat_seq <- "CCAAT.{100,150}ATG"
str_count(beta_replaced, caat_seq)
tata_seq <- "CTATAA.{50,100}ATG"
str_count(beta_replaced, tata_seq)
```

Find out if any caat_seq have tata_seq

```{r}
caat_seqs_all <- beta_replaced %>% 
  str_extract_all(caat_seq, simplify = TRUE)
str_which(caat_seqs_all, tata_seq)
promoter <- str_sub(caat_seqs_all[1])
promoter
str_length(promoter)
```

### Part 2: Ants

#### Import the data

import ants data.

```{r}
ants_import <- read_csv(here('data', 'ants.csv'))
ants_import
```

#### Wrangle the Data

tidy data

```{r}
ants_tidy <- ants_import %>% 
  pivot_longer(names_to = "species", values_to = "abundance", matches("[A-Z].+\\..+", ignore.case = FALSE)) %>% 
  select(year, species, abundance)
ants_tidy
```

Use mutate to change species names and then use mutate to turn year into ordered factor, group by year and species

```{r}
ants_mutated <- ants_tidy %>% 
  mutate(species = str_replace(species, "(^[A-Z][a-z]{0,3}).*\\.([a-z]{1,4}).*", "\\1\\2"))
ants_mutated
ants_year <- ants_mutated %>% 
  mutate(year = factor(year, levels = c(2010, 2011), ordered = TRUE))
ants_grouped <- ants_year %>% 
  group_by(year, species)
ants_summarized <- ants_grouped %>% 
  summarize('total abundance' = sum(abundance))
ants_summarized
```

#### Plot the Data

Make cleveland dot plot

```{r}
ants_summarized %>% 
  ggplot(aes(x = reorder(species, `total abundance`),
                         y = `total abundance`), group = species) +
  geom_line() +
  geom_point(mapping = aes(shape = year), size = 2) +
  coord_flip() +
  labs(x = NULL, y = 'Total Abundance', shape = 'Year') +
  theme_minimal()
```

### Part 3 :Featuring Phlox Flowers

#### Import and wrangle the first data set

```{r}
flower_import <- read_csv(here('data', 'flower_size.csv'))
flower_import
flower_wrangled <- flower_import %>% 
  filter(!is.na('Flower number')) %>% 
  select(species = 'Species', cor_length = `Corolla length (cm)`, throat_width = `Corolla width throat (cm)`) %>% 
  separate(species, into = c('genus', 'species'), extra = 'drop') %>% 
  group_by(genus, species) %>% 
  summarise(mean_length = mean(cor_length, na.rm = TRUE),
            mean_width = mean(throat_width, na.rm = TRUE),
            max_length = max(cor_length),
            max_width = max(throat_width)
            ,.groups = 'drop')
```

#### Import and wrangle the second data set

```{r}
pollinators_raw <- read_csv(here('data', 'flower_pollinators.csv'))
pollinators <- pollinators_raw %>% 
  select(-(Color | starts_with("Source"))) %>% 
  drop_na("Pollinator") %>% 
  separate(Species, into = c("genus", "species"), extra = 'drop') %>% 
  separate(Pollinator, into = "Syndrome", extra = 'drop')
pollinators
```

#### Join the Datasets









